A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2022-06-22 writeUp/supplementary_materials_methods.tex
added final submission

The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig. \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month, this assumption will likely be violated. To try to mitigate against such a risk, we aggregated the data across all dates to obtain a single series for each identifier, resulting in 201 such series. To obtain a reasonable level of accuracy on estimates, we removed all those individual (aggregated) series where fewer than 100 mosquitoes were collected. Finally, we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Fig. \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

@@

\subsection{Data collection for gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
To convert the estimates of lifespan in physiological age into chronological age requires a duration for the gonotrophic cycle. To determine this characteristic, we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle. A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'. The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}. Based on the abstracts of the resultant list of published studies, we then decided whether to search each article for estimates of the duration of the gonotrophic cycle. Overall 79 separate estimates of this parameter were found across 42 published articles (see Section \ref{sec:appendix_gonotrophicDurationStudyList} for a list of the studies included in our dataset). Along with information about the estimates, we also recorded study and series meta-data, including the location of the study, the method used for estimation, species and genus.

@@

Along with point estimates of the parameter we also collected information about the uncertainty in the estimates (if available). In many articles, the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately. If there was no disaggregation of durations between 1st and subsequent cycles, we assumed that each of these were the same. Raw estimates of gonotrophic cycle duration were obtained for species across three different genera (\textit{Aedes}, \textit{Anopheles} and \textit{Culex}), although there was a bias towards \textit{Anopheles} with $n=92$ estimates (Fig. \ref{fig:dissection_gonotrophicCycleRaw}).

>>>>

The data collection method resulted in 568 separate dissection series (see \verb|polovodova_full.csv|), across 73 published articles (see \verb|parity_papers_summary.csv|). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig. \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month, this assumption will likely be violated. To try to mitigate against such a risk, we aggregated the data across all dates to obtain a single series for each identifier, resulting in 201 such series (see \verb|polovodova_aggregated.csv|). To obtain a reasonable level of accuracy on estimates, we removed all those individual (aggregated) series where fewer than 100 mosquitoes were collected. Finally, we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Fig. \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

@@

\subsection{Data collection for gonotrophic cycle duration}\label{sec:dissection_gonotrophicData}
To convert the estimates of lifespan in physiological age into chronological age requires a duration for the gonotrophic cycle. To determine this characteristic, we conducted a meta-analysis of previously-published studies that estimate the duration of the gonotrophic cycle. A search of the literature using Google Scholar (\url{scholar.google.co.uk}) was performed using the search term: `gonotrophic cycle duration'. The list of articles was then supplemented with a list of references discussed by \cite{silver2007mosquito}. Based on the abstracts of the resultant list of published studies, we then decided whether to search each article for estimates of the duration of the gonotrophic cycle. Overall 79 separate estimates of this parameter were found across 42 published articles (see \verb|gonotrophic.csv|). Along with information about the estimates, we also recorded study and series meta-data, including the location of the study, the method used for estimation, species and genus.

@@

Along with point estimates of the parameter, we also collected information about the uncertainty in the estimates (if available). In many articles, the duration of the gonotrophic cycle was estimated separately for the first versus subsequent cycles, and these estimates were recorded separately. If there was no disaggregation of durations between 1st and subsequent cycles, we assumed that each of these were the same. Raw estimates of gonotrophic cycle duration were obtained for species across three different genera (\textit{Aedes}, \textit{Anopheles} and \textit{Culex}), although there was a bias towards \textit{Anopheles} with $n=92$ estimates (Fig. \ref{fig:dissection_gonotrophicCycleRaw}).
